### 2017-04-28 paper/jfp-submission/evaluation.tex
add a tldr results section

\paragraph{Benchmarks}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic and algebraic data types. % and expression evaluators.

@@

We answer the last question on a subset of the \ucsdbench dataset.
%
For each ill-typed program we identify the student's \emph{fix}
by searching for the first type-correct program that follows it in
the interaction trace.
%
We then use an expression-level \emph{diff}~\cite{Lempsink2009-xf} to
determine which sub-expressions changed between the ill-typed program
and the student's fix, and treat those expressions as the source of the
type error.

@@

\paragraph{Example: Recursion with Bad Base Case}
%
The function @sumList@ should add up
the elements of its input list.
%
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> (*@\hlSherrloc{hello}@*)
    | y::ys -> y + (*@\hlOcaml{sumList ys}@*)
\end{ecode}
%
Unfortunately, in the base case, it returns @[]@
instead of @0@.
%
\sherrloc\ blames the base case, and \ocaml\
assumes the base case is correct and blames
the \emph{recursive call} on line 3:
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
Both of the above are parts of the full story, which
is summarized by \toolname's trace showing
how @sumList [1; 2]@ gets stuck at @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
The trace clarifies (via the third step)
that the @[]@ results from the recursive call
\hbox{@sumList []@,} and shows how it is incompatible with
the subsequent \texttt{+} operation.

>>>>

In the sequel we present our experimental methodology (\S~\ref{sec:methodology})
and then answer the above questions.
%
However, for the impatient reader, we begin with a quick summary of our main results:

\paragraph{1. Most Type Errors Admit Witnesses}
Our prime result is that the vast majority of static type errors, around
85\%, do in fact admit a dynamic witness.
%
Further, \toolname efficiently synthesizes witnesses with its randomized search;
it can synthesize a witness for over 75\% of programs in under one second, \ie
fast enough to be integrated into the edit-compile-debug cycle.
%

\paragraph{2. Jump-Compressed Traces Are Small}
We find that our jump-compression heuristic effectively abstracts the
pedestrian details of computation, compressing the median trace with
14--15 single-step reductions to only 4 jumps.
%
Over 80\% of programs have a jump-compressed trace with at most 10
jumps, providing a bird's-eye view from which we can launch a more
in-depth investigation.

\paragraph{3. Witnesses Help Novices}
A witness should also help programmers \emph{understand} and
\emph{fix} type errors.
%
We use a set of ill-typed student programs to show that \toolname's
witnesses effectively demonstrate the runtime error that the type
system prevented.
%
Furthermore, we find, in a study of undergraduate students, that
\toolname's witnesses lead to more accurate diagnoses and fixes of type
errors than \ocaml's type error messages.

\paragraph{4. Witnesses Assign Blame}
Finally, we find that a simple heuristic allows us to use witnesses
to \emph{automatically} assign blame for type errors.
%
We treat the values inside the stuck term as \emph{sources} of typing
constraints and the stuck term itself as a \emph{sink}, producing
a small slice of the program that likely contains the error.
%
Using this heuristic, \toolname's witness are competitive with the
state-of-the-art type error localization tools \mycroft and \sherrloc.

\subsection{Methodology}
\label{sec:methodology}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic and algebraic data types. % and expression evaluators.

@@

We answer the last question on a subset of the \ucsdbench dataset.
%
For each ill-typed program in a student's interaction trace, we identify
the student's \emph{fix} by searching for the first type-correct program
that follows it in the trace.
%
We then use an expression-level \emph{diff}~\cite{Lempsink2009-xf} to
determine which sub-expressions changed between the ill-typed program
and the student's fix, and treat those expressions as the source of the
type error.

@@

\paragraph{Example: Recursion with Bad Base Case}
%
The function @sumList@ should add up
the elements of its input list.
%
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> (*@\hlSherrloc{[]}@*)
    | y::ys -> y + (*@\hlOcaml{sumList ys}@*)
\end{ecode}
%
Unfortunately, in the base case, it returns @[]@
instead of @0@.
%
\sherrloc\ blames the base case, and \ocaml\
assumes the base case is correct and blames
the \emph{recursive call} on line 3:
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
Both of the above are parts of the full story, which
is summarized by \toolname's trace showing
how @sumList [1; 2]@ gets stuck at @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
The trace clarifies (via the third step)
that the @[]@ results from the recursive call
\hbox{@sumList []@,} and shows how it is incompatible with
the subsequent \texttt{+} operation.

### 2017-04-28 paper/jfp-submission/locating.tex
todo
add a tldr results section

As discussed at the beginning of \S~\ref{sec:evaluation}, we recorded
each interaction of our students with the \ocaml top-level system.
%
In addition to collecting ill-typed programs, we collected
subsequent, fixed versions of the programs.
%
We identify fixed programs by searching for the first type-correct
program that follows the ill-typed program in the interaction trace.
%
We then use an expression-level \emph{diff}~\cite{Lempsink2009-xf} to
determine which sub-expressions changed between the ill-typed program
and the student's fix, and treat those expressions as the source of the
type error.

@@

>>>>

As discussed in \S~\ref{sec:methodology}, we recorded
each interaction of our students with the \ocaml top-level system.
%
In addition to collecting ill-typed programs, we collected
subsequent, fixed versions of the programs.
%
For each ill-typed program in a student's interaction trace, we identify
the student's fix by searching for the first type-correct program
that follows it in the trace.
%
We then use an expression-level \emph{diff}~\cite{Lempsink2009-xf} to
determine which sub-expressions changed between the ill-typed program
and the student's fix, and treat those expressions as the source of the
type error.

@@

\ES{TODO: threats to validity}
